Consumer extinctions constrain phenotypic evolution in the resulting food web

### 2019-08-30 analyses/reproduce_analyses.Rmd
tighten up curvature calculations and estimate long-term effects on evolvability

```{r Curvature, include=FALSE}
# Get mean estimates for directional selection gradients
foodweb_original_mean_sc.Diam <- filter(tidy_foodweb_grads, type == "Original", term == "sc.Diam")$mean
foodweb_original_mean_sc.log.Clutch <- filter(tidy_foodweb_grads, type == "Original", term == "sc.log.Clutch")$mean
foodweb_original_mean_sc.sqrt.Pref <- filter(tidy_foodweb_grads, type == "Original", term == "sc.sqrt.Pref")$mean

foodweb_extinction_mean_sc.Diam <- filter(tidy_foodweb_grads, type == "Extinction", term == "sc.Diam")$mean
foodweb_extinction_sc.log.Clutch <- filter(tidy_foodweb_grads, type == "Extinction", term == "sc.log.Clutch")$mean
foodweb_extinction_sc.sqrt.Pref <- filter(tidy_foodweb_grads, type == "Extinction", term == "sc.sqrt.Pref")$mean

@@

# Calculate curvatures for extinction food web
foodweb_extinction_raw_curvs <- foodweb_extinction_raw_gradients %>%
  transmute(`FoodwebExtinction:C.Diam` = `FoodwebExtinction:I(sc.Diam^2)` - foodweb_extinction_mean_sc.Diam*foodweb_extinction_mean_sc.Diam, #`FoodwebExtinction:sc.Diam`*`FoodwebExtinction:sc.Diam`,
         `FoodwebExtinction:C.Clutch` = `FoodwebExtinction:I(sc.log.Clutch^2)` - foodweb_extinction_sc.log.Clutch*foodweb_extinction_sc.log.Clutch, # `FoodwebExtinction:sc.log.Clutch`*`FoodwebExtinction:sc.log.Clutch`,
         `FoodwebExtinction:C.Pref` = `FoodwebExtinction:I(sc.sqrt.Pref^2)` - foodweb_extinction_sc.sqrt.Pref*foodweb_extinction_sc.sqrt.Pref,# `FoodwebExtinction:sc.sqrt.Pref`*`FoodwebExtinction:sc.sqrt.Pref`,
         `FoodwebExtinction:C.Diam:Clutch` = `FoodwebExtinction:sc.Diam:sc.log.Clutch` - foodweb_extinction_mean_sc.Diam*foodweb_extinction_sc.log.Clutch, # `FoodwebExtinction:sc.Diam`*`FoodwebExtinction:sc.log.Clutch`,
         `FoodwebExtinction:C.Diam:Pref` = `FoodwebExtinction:sc.Diam:sc.sqrt.Pref` - foodweb_extinction_mean_sc.Diam*foodweb_extinction_sc.sqrt.Pref,# `FoodwebExtinction:sc.Diam`*`FoodwebExtinction:sc.sqrt.Pref`,
         `FoodwebExtinction:C.Clutch:Pref` = `FoodwebExtinction:sc.log.Clutch:sc.sqrt.Pref` - foodweb_extinction_sc.log.Clutch*foodweb_extinction_sc.sqrt.Pref)# `FoodwebExtinction:sc.log.Clutch`*`FoodwebExtinction:sc.sqrt.Pref`)

@@

```{r mean curvature effects on random G-matrices}
# Generate random G-matrices
random_Gmats <- RandomMatrix(num.traits = 3, num.matrices = 10000, min.var = 0, max.var = 0.5)

@@

```{r Delta evolvability, fig.cap = "Change in evolvability for 10,000 random G-matrices"}
delta_evolvability_plot <- ggplot(random_evolvabilities, aes(x = Diff)) +
  geom_histogram() +
  geom_vline(xintercept = 0, linetype = "dotted") +
  xlab(expression(paste(Delta," evolvability (extinction - original)"))) +
  ylab("Count")

@@

>>>>

```{r Curvature}
# Get mean estimates for directional selection gradients
foodweb_original_mean_sc.Diam <- filter(tidy_foodweb_grads, type == "Original", term == "sc.Diam")$mean
foodweb_original_mean_sc.log.Clutch <- filter(tidy_foodweb_grads, type == "Original", term == "sc.log.Clutch")$mean
foodweb_original_mean_sc.sqrt.Pref <- filter(tidy_foodweb_grads, type == "Original", term == "sc.sqrt.Pref")$mean

foodweb_extinction_mean_sc.Diam <- filter(tidy_foodweb_grads, type == "Extinction", term == "sc.Diam")$mean
foodweb_extinction_mean_sc.log.Clutch <- filter(tidy_foodweb_grads, type == "Extinction", term == "sc.log.Clutch")$mean
foodweb_extinction_mean_sc.sqrt.Pref <- filter(tidy_foodweb_grads, type == "Extinction", term == "sc.sqrt.Pref")$mean

@@

# Calculate curvatures for extinction food web
foodweb_extinction_raw_curvs <- foodweb_extinction_raw_gradients %>%
  transmute(`FoodwebExtinction:C.Diam` = `FoodwebExtinction:I(sc.Diam^2)` - foodweb_extinction_mean_sc.Diam*foodweb_extinction_mean_sc.Diam, #`FoodwebExtinction:sc.Diam`*`FoodwebExtinction:sc.Diam`,
         `FoodwebExtinction:C.Clutch` = `FoodwebExtinction:I(sc.log.Clutch^2)` - foodweb_extinction_mean_sc.log.Clutch*foodweb_extinction_mean_sc.log.Clutch, # `FoodwebExtinction:sc.log.Clutch`*`FoodwebExtinction:sc.log.Clutch`,
         `FoodwebExtinction:C.Pref` = `FoodwebExtinction:I(sc.sqrt.Pref^2)` - foodweb_extinction_mean_sc.sqrt.Pref*foodweb_extinction_mean_sc.sqrt.Pref,# `FoodwebExtinction:sc.sqrt.Pref`*`FoodwebExtinction:sc.sqrt.Pref`,
         `FoodwebExtinction:C.Diam:Clutch` = `FoodwebExtinction:sc.Diam:sc.log.Clutch` - foodweb_extinction_mean_sc.Diam*foodweb_extinction_mean_sc.log.Clutch, # `FoodwebExtinction:sc.Diam`*`FoodwebExtinction:sc.log.Clutch`,
         `FoodwebExtinction:C.Diam:Pref` = `FoodwebExtinction:sc.Diam:sc.sqrt.Pref` - foodweb_extinction_mean_sc.Diam*foodweb_extinction_mean_sc.sqrt.Pref,# `FoodwebExtinction:sc.Diam`*`FoodwebExtinction:sc.sqrt.Pref`,
         `FoodwebExtinction:C.Clutch:Pref` = `FoodwebExtinction:sc.log.Clutch:sc.sqrt.Pref` - foodweb_extinction_mean_sc.log.Clutch*foodweb_extinction_mean_sc.sqrt.Pref)# `FoodwebExtinction:sc.log.Clutch`*`FoodwebExtinction:sc.sqrt.Pref`)

@@

```{r Mean curvature effects on random G-matrices, echo=TRUE}
## Generate random G-matrices
random_Gmats <- RandomMatrix(num.traits = 3, num.matrices = 10000, min.var = 0, max.var = 0.5)

@@

```{r Delta evolvability, fig.cap = "Change in evolvability for 10,000 random G-matrices using our best (mean) estimates of the curvature matrix."}
delta_evolvability_plot <- ggplot(random_evolvabilities, aes(x = Diff)) +
  geom_histogram() +
  geom_vline(xintercept = 0, linetype = "dotted") +
  xlab(expression(paste(Delta," evolvability (extinction - original)"))) +
  ylab("Count")

@@

```{r Long-term mean curvature effects on random G-matrices, echo=TRUE}
## Use same random G-matrices as above

# Note that original food web is same as before

# Long-term extinction
foodweb_long_extinction_mean_C.Diam <- filter(tidy_foodweb_long_curvs, type == "Extinction", term == "C.Diam")$mean
foodweb_long_extinction_mean_C.Clutch <- filter(tidy_foodweb_long_curvs, type == "Extinction", term == "C.Clutch")$mean
foodweb_long_extinction_mean_C.Pref <- filter(tidy_foodweb_long_curvs, type == "Extinction", term == "C.Pref")$mean
foodweb_long_extinction_mean_C.Diam.Clutch <- filter(tidy_foodweb_long_curvs, type == "Extinction", term == "C.Diam:Clutch")$mean
foodweb_long_extinction_mean_C.Diam.Pref <- filter(tidy_foodweb_long_curvs, type == "Extinction", term == "C.Diam:Pref")$mean
foodweb_long_extinction_mean_C.Clutch.Pref <- filter(tidy_foodweb_long_curvs, type == "Extinction", term == "C.Clutch:Pref")$mean

long_mean_curv_extinction <- matrix(c(foodweb_long_extinction_mean_C.Diam, foodweb_long_extinction_mean_C.Diam.Clutch, foodweb_long_extinction_mean_C.Diam.Pref,
                               foodweb_long_extinction_mean_C.Diam.Clutch, foodweb_long_extinction_mean_C.Clutch, foodweb_long_extinction_mean_C.Clutch.Pref,
                               foodweb_long_extinction_mean_C.Diam.Pref, foodweb_long_extinction_mean_C.Clutch.Pref, foodweb_long_extinction_mean_C.Pref),
                             byrow = T, ncol = 3)

## Functions to calculate effect of mean curvature on random G-matrices
curv_long_extinction_random_matrices_evolvabilities <- function(x){
  new_Gmat <- x + x %*% long_mean_curv_extinction %*% x
  MeanMatrixStatistics(new_Gmat, iterations = 1000, parallel = TRUE)["evolvability"]
}

long_extinction_evolvability_list <- plyr::ldply(lapply(X = random_Gmats, FUN = curv_long_extinction_random_matrices_evolvabilities)) %>% 
  select(Extinction = evolvability)

long_random_evolvabilities <- bind_cols(original_evolvability_list, long_extinction_evolvability_list) %>%
  mutate(Diff = Extinction - Original)
```

Over the long-term (after egg parasitoids recover and impose similar fitness impact), we see that for our best estimates of the curvature matrix, that the extinction food web tends to decrease evolvability (`r sum(long_random_evolvabilities$Diff < 0) / length(long_random_evolvabilities$Diff)`) of the gall midge population.

```{r Long-term delta evolvability, fig.cap = "Change in evolvability for 10,000 random G-matrices using our best (mean) estimates of the curvature matrix over the long term."}
long_delta_evolvability_plot <- ggplot(long_random_evolvabilities, aes(x = Diff)) +
  geom_histogram() +
  geom_vline(xintercept = 0, linetype = "dotted") +
  xlab(expression(paste(Delta," evolvability (extinction - original)"))) +
  ylab("Count")

long_delta_evolvability_plot
```
